---
title: "Useful R code"
author: "Nils Hoyer"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Exploratory data analysis
## Table functions
The table function summarizes a table

```{r}
Gender <- c("Female","Female","Male","Male")
Restaurant <- c("Yes","No","Yes","No")
Count <- c(220, 780, 400, 600)
DiningSurvey <- data.frame(Gender, Restaurant, Count)
DiningSurvey
table(DiningSurvey)
```

Only include the values >650
```{r}
table(DiningSurvey$Count>650)

```

Also include missing 
```{r}
DiningSurvey$Restaurant <- c("Yes", "No", "Yes", NA)
table(DiningSurvey$Restaurant, useNA="always")
```

Exclude males
```{r}
table(DiningSurvey$Gender, exclude = "Male")
```

Use margin.table to summarize rows or collumns
```{r}
RentalUnits <- matrix(c(45,37,34,10,15,12,24,18,19),ncol=3,byrow=TRUE)
colnames(RentalUnits) <- c("Section1","Section2","Section3")
rownames(RentalUnits) <- c("Rented","Vacant","Reserved")
RentalUnits <- as.table(RentalUnits)
#Showing sums of rows
margin.table(RentalUnits, 1)
#Showing sums of columns
margin.table(RentalUnits, 2)
```

Use prop.table to show the percentages (proportions)
```{r}
#Cell percentages
prop.table(RentalUnits)
#Row percentages
prop.table(RentalUnits, 1)
#Columns percentages
prop.table(RentalUnits, 2)
```

Show the table in a pretty format
```{r}
ftable(RentalUnits)

```

Perform a Chi-squared test on the table (what is going on???)
```{r}
summary(RentalUnits)
```

Use as.data.frame to list frequencies
```{r}
as.data.frame(RentalUnits)
```

Use addmargins to add an extra row of sums or column of sums
```{r}
addmargins(RentalUnits)
#Can be used on prop.table
addmargins(prop.table(RentalUnits))

#Can be used only on columns
addmargins(prop.table(RentalUnits,1))


#Can be used only on rows
addmargins(prop.table(RentalUnits,2))
```



# Regression analysis
## Simple linear regression
Example from coursera course
```{r}
#Open library UsingR which contains the diamonds dataset
library(UsingR)
data("diamond")

#Start with a plot of price against mass (in carats). Add a fitted line 
plot(diamond$carat, diamond$price,
        xlab="Mass (carats)",
        ylab="Price (dollar)",
        bg="lightblue",
        col="black", cex=1.1, pch=21, frame=FALSE)
abline(lm(price~carat, data=diamond), lwd=2)

# Assign the coefficients from the linear regression to a "fit" variable and call its coefficients
fit <- lm(price~carat, data=diamond)
coef(fit)

#Center the data by subtracting the mean. this gives the intercept at the average x (the average size of diamond)
fit2 <- lm(price~I(carat - mean(carat)), data = diamond)
coef(fit2)

#Now we try to fit some values of x
newx <- c(0.16, 0.27, 0.34)
predict(fit, newdata = data.frame(carat = newx))

#Residuals. First create a regression
y <- diamond$price; x <- diamond$carat;n <- length(y)
fit <- lm(y~x)

#Call the sigma from summary which gives the residual variation
summary(fit)$sigma

#Call all the important coefficients
summary(fit)$coefficients

#Getting a confidence interval for the intercept, we call the appropriate cells in the coefficients table
sumCoef <- summary(fit)$coefficients
sumCoef[1,1] + c(-1,1) * qt(.975, df=fit$df) * sumCoef[1,2]

#Getting a confidence interval for the slope, we call the appropriate cells in the coefficients table
sumCoef <- summary(fit)$coefficients
sumCoef[2,1] + c(-1,1) * qt(.975, df=fit$df) * sumCoef[2,2]

```

